---
title: "Assignment5_Code"
author: "Stephanie Rosen"
date: "12/12/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction

# Description of Dataset



# Data Acquisition, Exploration, Filtering, and Quality Control

The packages we will be using are tidyverse, 

```{r Loading required packages, echo=FALSE}

library(tidyverse)

```


We require two datasets for this project: (1) Community data for the wetlands and (2) Genetic data from NCBI 

```{r community data acquistion, exploration, filtering and quality control}

#import dataset
df_Odonata <- PEI_NP_Wetlands_Odonata_2010_2018_data

names(df_Odonata)
dim(df_Odonata)

#total number of unique species?
#73
#but lots of errors with species names. We will remove NA values, any species names that do not follow proper scientific naming procedures, and also any where the species level is not indicated. 


#remove chunks of text in brackets but still have full species names
df_Odonata$Scientific.name <- str_remove(string = df_Odonata$Scientific.name, "\\(.+\\)" )

#There are some species that have a third unecessary word in their name. We will remove that too.
#df_Odonata[str_detect(string = df_Odonata$Scientific.name, "^[A-z]+ [A-z]+ [A-z]+$"),]

df_Odonata$Scientific.name <- str_remove(string = df_Odonata$Scientific.name, "Walker" )
df_Odonata$Scientific.name <- str_remove(string = df_Odonata$Scientific.name, "Say" )

#remove first row (French header), scientific names with NA values, irrelevant columns, and any names that do not follow. 
df_Odonata <- df_Odonata[-1,] %>%
  filter(!is.na(Scientific.name)) %>%
  select(!exuvia.observer.number & !Observer..order.number & !note) %>%
  filter(str_detect(Scientific.name, "^[A-z]+ [A-z]+$")) %>%
  filter(!str_detect(Scientific.name, "NOT ODONATA"))

unique(df_Odonata$Scientific.name)
unique(PEI_NP_Wetlands_Odonata_2010_2018_data$Scientific.name)

#Now that we have clean scientific names, let's look at the different ponds we have:
unique(df_Odonata$Survey.site)

```

Now we have clean scientific name data for our different communities. Let's explore the data

```{r data exploration}



```




# Main Software Tools Description

# Main Analysis


```{r pressure, echo=FALSE}


```


#Results and Discussion

#Acknowledgements

#References